On the analytic structure of the linear response susceptibility in open systems

### 2018-03-20 AnalyticStructure_main.tex
update

The subject of the present investigation concerns with the study of isolated molecular-like systems subjected to a time-dependent perturbation in the linear response regime.

The ground state properties of these systems are described within the standard DFT approach in which (valence) electrons are distributed on the Kohn-Sham orbitals. The stability of
the ground state and the localized nature of the system ensures that the KS description of occupied states is represented by a discrete set of wave functions that becomes arbitrary 
small far apart from a compact region around the system's atoms. This behavior indicates that a (finite) basis set will provide a reliable description of the ground state properties 
if it is able to sample a computational domain which contains the natural sizes of the occupied orbitals\cite{boffi2016}. %Results discussed in \cite{boffi2016} confirm this statement.

The same arguments cannot be rigidly extended to the case in which an external time-dependent field starts acting on the system. Indeed, the perturbation drives the system out 
from the equilibrium and the electrons sample a wider region of the state space, that will contains orbitals of both negative (bound) and positive (unbound) energies. Unbound orbitals 
form a continuum of states in the true system with open boundaries, and the introduction of a finite computational domain has the effects of introducing a fictitious level discretization
\cite{boffi2016}.

@@

The perturbation induces a modification on the ground state density and, choosing the
eigenstates of the unperturbed KS Hamiltonian as a natural basis, one has to include also the virtual (unoccupied) states to described the perturbed density. Virtual states of
positive energy constitute a continuum...

Sebbene non ci sia necessità di usare gli autostati di KS per campionare lo spazio ortogonale a quello in cui si esprime la densità imperturbata, questo argomento solleva un potenziale
problema: quale è la scelta di una basis set idonea per costruire la risposta lineare in modo affidabile?

@@

State something about the assessment of physical reliability for a ground state calculation. Some comment on the setup independence of the results. Introduce the correspondence
between localization properties and observable behavior or physicality. State the, at the present days, similar consideration for the results of linear response are still 
missing...

>>>>

The subject of the present investigation concerns with the study of isolated molecular-like systems subjected to a time-dependent perturbation. % in the linear response regime.
This problem can be tackled within the contest of the time-dependent density functional theory (TD-DFT)\cite{runge1984,onida2002}, that allows one to establish the time-dependent 
Kohn-Sham (KS) equations that determine the evolution of the one-particle properties of the system with respect to time. We focus our attention to the linear response regime (LR) 
in which the theory can be conveniently expressed in the frequency domain and the time-dependent KS equation are reformulated as a quantum Liouville equation for the response 
density operator. Different computational approaches have been developed in literature\cite{baroni2008,hubener2014,brabec2015}, principally with the aim of producing efficient algorithms 
for the evaluation of the physical observables. These methods differ both for the procedure adopted to solve the Liouville equation (explicit diagonalization, self-consistent Sternheimer 
equation, Lanczos chain-like procedure, etc.) and both for the choice of the basis set employed to perform the calculations. Nonetheless, in all the cases one is a dealing with a finite 
basis set that, unavoidably, gives rise to a limited computational domain. It is then meaningful to try to understand under which extent and for which type of observables the artificial 
truncation of the natural open boundaries condition could produce observable effects on the physical quantities computed in the LR.

A preliminary answer of this question can be formulated by considering results \cite{boffi2016}, that establish a clear correspondence between the orbital localization character, and 
the independence of its properties (\emph{e.g.} energy) with respect to the basis set. It is shown that (localized) orbitals with a bound-state behavior in the long range 
constitute a genuine discrete set and a finite basis is able to provide a reliable description of these states if the sampled domain contains their natural sizes. On the contrary, 
unbound (delocalized) orbitals form a continuum of states in the true system with open boundaries. For these class of objects the presence finite computational domain has a deeper 
impact since introduces a fictitious discretization of the energy levels. Furthermore, all the energies of the unbound orbitals scale to zero as long as the computational domain is 
increased.

First of all it is useful to discuss the property of the ground state under this perspective. 
Indeed, it is well known that the stability of the ground state and the localized nature of a molecular-like system ensures that the KS description of occupied states is represented 
by a discrete set of wave functions that approaches to zero far apart from a compact region around the system's atoms. This behavior indicates that a (finite) basis set will provide 
a reliable description of the ground state properties if it is able to sample a computational domain which contains the natural sizes of the occupied orbitals\cite{boffi2016}.

% A preliminary answer of this question can be formulated by considering results derived in \cite{boffi2016}, concerning the basis dependence of occupied and virtual molecular orbitals.   
% First of all it is useful to discuss the property of the ground state under this perspective. 
% Indeed, it is well known that the stability of the ground state and the localized nature of a molecular-like system ensures that the KS description of occupied states is represented 
% by a discrete set of wave functions that approaches to zero far apart from a compact region around the system's atoms. This behavior indicates that a (finite) basis set will provide 
% a reliable description of the ground state properties 
% if it is able to sample a computational domain which contains the natural sizes of the occupied orbitals\cite{boffi2016}.

The same arguments cannot be rigidly extended to the case in which an external time-dependent field starts acting on the system. Indeed, the perturbation drives the system out 
from the equilibrium and the electrons sample a wider region of the state space, that will contains orbitals of both negative (bound) and positive (unbound) energies. 
% Unbound orbitals 
% form a continuum of states in the true system with open boundaries, and the introduction of a finite computational domain has the effects of introducing a fictitious level discretization
% \cite{boffi2016}.

@@

% A preliminary answer of this question can be formulated by considering results derived in \cite{boffi2016}, concerning the basis dependence of occupied and virtual molecular orbitals.   
% Indeed, the ground state properties of molecular-like systems are described within the standard DFT approach in which electrons are distributed on the Kohn-Sham orbitals. The stability of
% the ground state and the localized nature of the system ensures that the KS description of occupied states is represented by a discrete set of wave functions that approaches to zero  
% far apart from a compact region around the system's atoms. This behavior indicates that a (finite) basis set will provide a reliable description of the ground state properties 
% if it is able to sample a computational domain which contains the natural sizes of the occupied orbitals\cite{boffi2016}.

% The perturbation induces a modification on the ground state density and, choosing the
% eigenstates of the unperturbed KS Hamiltonian as a natural basis, one has to include also the virtual (unoccupied) states to described the perturbed density. Virtual states of
% positive energy constitute a continuum...

% Sebbene non ci sia necessità di usare gli autostati di KS per campionare lo spazio ortogonale a quello in cui si esprime la densità imperturbata, questo argomento solleva un potenziale
% problema: quale è la scelta di una basis set idonea per costruire la risposta lineare in modo affidabile?

@@

% State something about the assessment of physical reliability for a ground state calculation. Some comment on the setup independence of the results. Introduce the correspondence
% between localization properties and observable behavior or physicality. State the, at the present days, similar consideration for the results of linear response are still 
% missing...
